---
title: "06_1_calculate_values_metafeatures"
author: "Bozeng_Tang"
date: "04/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r }
library(Seurat)
library(harmony)
library(SeuratWrappers)
library(dplyr)
library(tidyverse)

# reading the object from previous section
load("~/Desktop/Single_Cell/processed/NEW_bt1_to_6/06.1_manipulate_genes_metafeatures/saved_features.RData")
# before that I need to seperate the inter_list,
# it has to be manual, but we have see a lot of things.

DefaultAssay(SCT_result)<-"SCT"

calium_list<-inter_list%>%
  filter(Source=="calcium")

NLR_list<-inter_list%>%
  filter(Source=="NLR")

RLK_list<-inter_list%>%
  filter(Source=="RLK")

RLP_list<-inter_list%>%
  filter(Source=="RLP")

DAMPs_list<-inter_list%>%
  filter(Source=="MAMPs")

SA_list<-inter_list%>%
  filter(Source=="Salicylic Acid")

WRKY_list<-inter_list%>%
  filter(Source=="WYKY")

ROS_list<-inter_list%>%
  filter(Source=="ROS")

defense_list<-inter_list%>%
  filter(Source=="defense")

JA_list<-inter_list%>%
  filter(Source=="JA")

MAPK_list<-inter_list%>%
  filter(Source=="MAPk")

small_list<-inter_list%>%
  filter(Source=="small_RNA")

PAMPs_list<-inter_list%>%
  filter(Source=="PAMPs")

```

# I want to apply the same rational for the cell type markers, and then we can calculate the full metafeatures
```{r}
cell_types<-read.csv("raw_data/final_marker.csv")

indata_cell_markers<-cell_types%>%
  filter(gene%in%rownames(SCT_result))

M_marker_list<-indata_cell_markers%>%
  filter(Cell_type=="leaf mesophyll" | Cell_type=="leaf")


E_marker_list<-indata_cell_markers%>%
  filter(Cell_type=="leaf epidermis" | Cell_type=="leaf pavement cell")

G_marker_list<-indata_cell_markers%>%
  filter(Cell_type=="leaf guard cell" | Cell_type=="leaf stomatal complex"| Cell_type=="leaf stomatal pore")
 
V_marker_list<-indata_cell_markers%>%
  filter(Cell_type=="vascular leaf" )

P_marker_list<-indata_cell_markers%>%
  filter(Cell_type=="vascular leaf"| Cell_type=="phloem" )

B_marker_list<-indata_cell_markers%>%
  filter(Cell_type=="vascular leaf" | Cell_type=="bundle sheath" )
 

```

## now, let us calcualte metafeature for each cell, as it would be interesting

```{r}

# 

SCT_result <- MetaFeature( object = SCT_result, features = calium_list$gene_name,
                           meta.name = 'calium_meta')

SCT_result <- MetaFeature( object = SCT_result, features = NLR_list$gene_name,
                           meta.name = 'NLR_meta')

SCT_result <- MetaFeature( object = SCT_result, features = RLK_list$gene_name,
                           meta.name = 'RLK_meta')

SCT_result <- MetaFeature( object = SCT_result, features = RLP_list$gene_name,
                           meta.name = 'RLP_meta')

SCT_result <- MetaFeature( object = SCT_result, features = DAMPs_list$gene_name,
                           meta.name = 'DAMPs_meta')

SCT_result <- MetaFeature( object = SCT_result, features = SA_list$gene_name,
                           meta.name = 'SA_meta')

SCT_result <- MetaFeature( object = SCT_result, features = WRKY_list$gene_name,
                           meta.name = 'WRKY_meta')

SCT_result <- MetaFeature( object = SCT_result, features = PAMPs_list$gene_name,
                           meta.name = 'PAMPs_meta')

SCT_result <- MetaFeature( object = SCT_result, features = ROS_list$gene_name,
                           meta.name = 'ROS_meta')

SCT_result <- MetaFeature( object = SCT_result, features = defense_list$gene_name,
                           meta.name = 'defense_meta')

SCT_result <- MetaFeature( object = SCT_result, features = JA_list$gene_name,
                           meta.name = 'JA_meta')

SCT_result <- MetaFeature( object = SCT_result, features = MAPK_list$gene_name,
                           meta.name = 'MAPK_meta')

SCT_result <- MetaFeature( object = SCT_result, features = known_list$gene_name,
                           meta.name = 'known_meta')

SCT_result <- MetaFeature( object = SCT_result, features = laser_list$gene_name,
                           meta.name = 'laser_meta')

SCT_result <- MetaFeature( object = SCT_result, features = AP_list$gene_name,
                           meta.name = 'AP_meta')

SCT_result <- MetaFeature( object = SCT_result, features = BIO_list$gene_name,
                           meta.name = 'BIO_meta')

SCT_result <- MetaFeature( object = SCT_result, features = PC_list$gene_name,
                           meta.name = 'pc_meta')

# now, lets add the marker genes
SCT_result <- MetaFeature( object = SCT_result, features = M_marker_list$gene,
                           meta.name = 'mesophyll_meta')

SCT_result <- MetaFeature( object = SCT_result, features = E_marker_list$gene,
                           meta.name = 'Epidermis_meta')

SCT_result <- MetaFeature( object = SCT_result, features = G_marker_list$gene,
                           meta.name = 'Guard_meta')

SCT_result <- MetaFeature( object = SCT_result, features = V_marker_list$gene,
                           meta.name = 'vasculature_meta')

SCT_result <- MetaFeature( object = SCT_result, features = P_marker_list$gene,
                           meta.name = 'Plohem_meta')

SCT_result <- MetaFeature( object = SCT_result, features = B_marker_list$gene,
                           meta.name = 'bundle_meta')

```

## AddmoduleScore
```{r}
## need to transform the dataframe column, into list
calium_score_list<-list(c(calium_list$gene_name))
NLR_score_list<-list(c(NLR_list$gene_name))
RLK_score_list<-list(c(RLK_list$gene_name))
RLP_score_list<-list(c(RLP_list$gene_name))
DAMPs_score_list<-list(c(DAMPs_list$gene_name))
SA_score_list<-list(c(SA_list$gene_name))
WRKY_score_list<-list(c(WRKY_list$gene_name))
ROS_score_list<-list(c(ROS_list$gene_name))
defense_score_list<-list(c(defense_list$gene_name))
JA_score_list<-list(c(JA_list$gene_name))
MAPK_score_list<-list(c(MAPK_list$gene_name))
PAMPs_score_list<-list(c(PAMPs_list$gene_name))
known_score_list<-list(c(known_list$gene_name))
laser_score_list<-list(c(laser_list$gene_name))
AP_score_list<-list(c(AP_list$gene_name))
Bio_score_list<-list(c(BIO_list$gene_name))
PC_score_list<-list(c(PC_list$gene_name))

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = calium_score_list, name = 'calium_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = NLR_score_list, name = 'NLR_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = RLK_score_list, name = 'RLK_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = RLP_score_list, name = 'RLP_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = DAMPs_score_list, name = 'DAMPs_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = SA_score_list, name = 'SA_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = WRKY_score_list, name = 'WRKY_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = ROS_score_list, name = 'ROS_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = defense_score_list, name = 'defense_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = JA_score_list, name = 'JA_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = MAPK_score_list, name = 'MAPK_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = PAMPs_score_list, name = 'PAMPs_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = known_score_list, name = 'known_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = laser_score_list, name = 'laser_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = AP_score_list, name = 'AP_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = Bio_score_list, name = 'Bio_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = PC_score_list, name = 'PC_score')
```

# add HR, PTI, 
```{r}
PTI_core<-read.csv("raw_data/defense_genes/PTI/core_PTI_40.csv")
PTI_up<-read.csv("raw_data/defense_genes/PTI/PTI_700up.csv")
PTI_down<-read.csv("raw_data/defense_genes/PTI/PTI_down_100.csv")

HR_in<-read.csv("raw_data/defense_genes/HR_markers/HR_in_specific.csv")
HR_out<-read.csv("raw_data/defense_genes/HR_markers/out_HR_specific.csv")
HR_inandout<-read.csv("raw_data/defense_genes/HR_markers/In_and_Out_HR.csv")

# list for core
PTI_core_list<-inner_join(PTI_core, feature_list, by=c("core_PTI_40" ="locus"))%>%
  filter(gene_name%in%rownames(SCT_result))
# list fore core score
PTI_core_score_list<-list(c(PTI_core_list$gene_name))

# list up PTI
PTI_up_list<-inner_join(PTI_up, feature_list, by=c("PTI_700_up" ="locus"))%>%
  filter(gene_name%in%rownames(SCT_result))

#list up socre pti
PTI_up_score_list<-list(c(PTI_up_list$gene_name))

PTI_down_list<-inner_join(PTI_down, feature_list, by=c("PTI_down_100" ="locus"))%>%
  filter(gene_name%in%rownames(SCT_result))

HR_in_list<-inner_join(HR_in, feature_list, by=c("HR_in_specific" ="locus"))%>%
  filter(gene_name%in%rownames(SCT_result))

HR_out_list<-inner_join(HR_out, feature_list, by=c("out_HR_specific" ="locus"))%>%
  filter(gene_name%in%rownames(SCT_result))

HR_inandout_list<-inner_join(HR_inandout, feature_list, by=c("IN_and_out_HR" ="locus"))%>%
  filter(gene_name%in%rownames(SCT_result))

# Meatafeature HR, PTI
SCT_result <- MetaFeature( object = SCT_result, features = PTI_core_list$gene_name,
                           meta.name = 'PTI_core_meta')

SCT_result <- MetaFeature( object = SCT_result, features = PTI_up_list$gene_name,
                           meta.name = 'PTI_up_meta')

SCT_result <- MetaFeature( object = SCT_result, features = PTI_down_list$gene_name,
                           meta.name = 'PTI_down_meta')

SCT_result <- MetaFeature( object = SCT_result, features = HR_in_list$gene_name,
                           meta.name = 'HR_in_meta')

SCT_result <- MetaFeature( object = SCT_result, features = HR_out_list$gene_name,
                           meta.name = 'HR_out_meta')

SCT_result <- MetaFeature( object = SCT_result, features = HR_inandout_list$gene_name,
                           meta.name = 'HR_inandout_meta')

```

# add metascore for cell markers, HR, and PTI genes
```{r}
mark_M_score_list<-list(c(M_marker_list$gene))

mark_E_score_list<-list(c(E_marker_list$gene))

mark_G_score_list<-list(c(G_marker_list$gene))

mark_V_score_list<-list(c(V_marker_list$gene))

mark_P_score_list<-list(c(P_marker_list$gene))

mark_B_score_list<-list(c(B_marker_list$gene))

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = mark_M_score_list, name = 'Mesophyll_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = mark_E_score_list, name = 'Epidermis_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = mark_G_score_list, name = 'guardcell_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = mark_V_score_list, name = 'vasculature_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = mark_P_score_list, name = 'phloem_score')

SCT_result <- AddModuleScore( object = SCT_result,   ctrl = 5,
  features = mark_B_score_list, name = 'bunl_shealth_score')
```

# we add aggregation values for plotting
# we save the SCT dataset and this Rdata, for future plotting
# calculate Agggre and introduce into Seurat
# the output of AGG would be a matrix of each gene, with the combination of clusters, and treatment.
# it would be useful to produce heatmap, to illustrate the detailed inforamtion of gene in each clusters
```{r}
# define the clusters as the ident
SCT_result<-SetIdent( SCT_result, value = "seurat_clusters" )
DefaultAssay(SCT_result)<-"SCT"

# produce aggree for all the lists:

cal_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = calium_list$gene_name))

NLR_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = NLR_list$gene_name))

RLK_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = RLK_list$gene_name))

RLP_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = RLP_list$gene_name))

DAMPs_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = DAMPs_list$gene_name))

SA_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = SA_list$gene_name))

WRKY_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = WRKY_list$gene_name))

ROS_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = ROS_list$gene_name))

defense_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = defense_list$gene_name))

JA_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = JA_list$gene_name))

MAPK_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = MAPK_list$gene_name))


PAMPs_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = PAMPs_list$gene_name))

known_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = known_list$gene_name))

laser_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = laser_list$gene_name))

AP_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = AP_list$gene_name))

BIO_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = BIO_list$gene_name))

PC_ag<-as.data.frame( AggregateExpression(object = SCT_result, 
                    group.by=c("orig.ident","pathogen_Cells"), # here we create the combination of 
                    features = PC_list$gene_name))
```


```{r}
saveRDS(SCT_result, file='processed/NEW_bt1_to_6/07.1_create_values/added_metafeatures')

save.image("~/Desktop/Single_Cell/processed/NEW_bt1_to_6/07.1_create_values/meta_aggre_all.RData")
```